Supplementary Materials for ``Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal"

### 2025-12-10 submissions/TilingUploadROPP/Tiling_Project/SupplementaryPR.tex
started creating referee response

%\documentclass{article}
%\documentclass[
%%preprint,
%reprint,onecolumn,
%superscriptaddress,
%groupedaddress,
%%unsortedaddress,
%%runinaddress,
%%frontmatterverbose, 
%%preprint,
%%preprintnumbers,
%%nofootinbib,
%%nobibnotes,
%%bibnotes,
%%aps,superscriptaddress,
%aps,pra,
%%rmp,
%%prstab,
%%prstper,
%%floatfix,
%10pt
%]{revtex4-2}
\documentclass[%
 reprint,onecolumn,
superscriptaddress,
groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
aps,superscriptaddress,
pra,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}
%\usepackage{amsmath,amssymb,bm,graphicx,braket}
\usepackage[margin=1.5cm]{geometry}
%\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{graphicx,braket,color,bm,amsmath,amssymb,braket,placeins}
% \graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}
\usepackage{natbib}

@@

\makeatletter
\apptocmd{\thebibliography}{\global\c@NAT@ctr 84\relax}{}{}
\makeatother
\begin{document}
\title{Supplementary Materials for ``Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal"}

\author{Abhirup Mukherjee}
\email{am18ip014@iiserkol.ac.in}
\affiliation{Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}
\author{S. R. Hassan}
\email{shassan@imsc.res.in}
\affiliation{The Institute of Mathematical Sciences, HBNI, C.I.T. Campus, Chennai 600 113, India}
\author{Anamitra Mukherjee}
\email{anamitra@niser.ac.in}
\affiliation{School of Physical Sciences, National Institute of Science, Education and Research, HBNI, Jatni 752050, India}
\author{N. S. Vidhyadhiraja}
\email{raja@jncasr.ac.in}
\affiliation{Theoretical Sciences Unit, Jawaharlal Nehru Center for Advanced Scientific Research, Jakkur, Bengaluru 560064, India}
\author{A. Taraphder}
\email{arghya@phy.iitkgp.ac.in}
\affiliation{Department of Physics, Indian Institute of Technology Kharagpur, Kharagpur 721302, India}
\author{Siddhartha Lal}
\email{slal@iiserkol.ac.in}
\affiliation{Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}

@@

\section{Derivation of unitary RG equations for the lattice-embedded impurity model}

\subsection{The unitary renormalisation group method}
In order to obtain the various low-energy phases of our impurity model, we perform a scaling analysis of the associated Hamiltonian using the recently developed unitary renormalisation group (URG) method ~[24,42,43].
%\cite{anirbanurg1,anirbanurg2}. 
The method has been applied successfully on a wide variety of problems of correlated fermions~\cite{santanukagome,1dhubjhep1,anirbanmott1,siddharthacpi,anirban_kondo,Patra_2023}. The method proceeds by resolving quantum fluctuations in high-energy degrees of freedom, leading to a low-energy Hamiltonian with renormalised couplings and new emergent degrees of freedom. Typically, for a system with Fermi energy \(\epsilon_F\) and bandwidth \(E_N\), the sequence of isoenergetic shells \(\left\{E_{(j)}\right\}, E_{j}\in \left[E_0, E_N\right] \) define the states whose quantum fluctuations we sequentially resolve. The momentum states lying on shells \(E_N\) that are far away from the Fermi surface comprise the UV states, while those on shells near the Fermi surface comprise the IR states.

As a result of the URG transformations, the Hamiltonian \(H_{(j)}\) at a given RG step \(j\) involves scattering processes between the \(k-\)states that have energies lower than \(D_{(j+1)}\). The unitary transformation \(U_{(j)}\) is then defined so as to remove the number fluctuations of the currently most energetic set of states \(D_{(j)}\)~[42,43]:
%\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	H_{(j-1)} = U_{(j)} H_{(j)} U^\dagger_{(j)}~, \text{such that} ~\left[H_{(j-1)}, \hat n_{j}\right] =0~.
\end{eqnarray}
The eigenvalue of $\hat{n}_{j}$ has, thus, been rendered an integral of motion (IOM) under the RG transformation.

The unitary transformations can be expressed in terms of a generator \(\eta_{(j)}\) that has fermionic algebra~[42,43]:
%\cite{anirbanurg1,anirbanurg2}:
\begin{eqnarray}
	\label{unitary}
	U_{(j)} = \frac{1}{\sqrt 2}\left(1 + \eta_{(j)} - \eta_{(j)}^\dagger\right)~,~ \quad\left\{ \eta_{(j)},\eta_{(j)}^\dagger \right\} = 1~,
\end{eqnarray}
where \(\left\{\cdot\right\}\) is the anticommutator. The unitary operator \(U_{(j)}\) that appears in Eq.~\eqref{unitary} can be cast into the well-known general form \(U = e^\mathcal{S}, \mathcal{S} = \frac{\pi}{4}\left( \eta^\dagger_{(j)} - \eta_{(j)} \right)\) that a unitary operator can take, defined by an anti-Hermitian operator \(\mathcal{S}\). The generator \(\eta_{(j)}\) is given by the expression~[42,43]:
%\cite{anirbanurg1,anirbanurg2}
\begin{eqnarray}
	\eta^\dagger_{(j)} = \frac{1}{\hat \omega_{(j)} - \text{Tr}\left(H_{(j)} \hat n_{j}\right) } c^\dagger_{j} \text{Tr}\left(H_{(j)}c_{j}\right)~.
\end{eqnarray}
The operators \(\eta_{(j)},\eta^\dagger_{(j)}\) behave as the many-particle analogues of the single-particle field operators \(c_j,c^\dagger_j\) - they change the occupation number of the single-particle Fock space \(\ket{n_j}\).  The important operator \(\hat \omega_{(j)}\) originates from the quantum fluctuations that exist in the problem because of the non-commutation of the kinetic energy terms and the interaction terms in the Hamiltonian:
\begin{eqnarray}
	\hat \omega_{(j)} = H_{(j-1)} - H^i_{(j)}~.
	\label{omega}
\end{eqnarray}
\(H^i_{(j)}\) is the part of \(H_{(j)}\) that commutes with \(\hat n_j\) but does {\it not} commute with at least one \(\hat n_l\) for \(l < j\). The RG flow continues up to energy \(D^*\), where a fixed point is reached from the vanishing of the RG function. 
Detailed comparisons of the URG with other methods (e.g., the functional RG, spectrum bifurcation RG etc.) can be found in Refs.~[42,43].
%\cite{anirbanmott1,anirbanurg1}.

\subsection{RG scheme}
At any given step \(j\) of the RG procedure, we decouple the states \(\left\{ {\bf q} \right\} \) on the isoenergetic surface of energy \(\varepsilon_j\). The diagonal Hamiltonian \(H_D\) for this step consists of all terms that do not change the occupancy of the states \(\left\{{\bf q}\right\}\):
\begin{equation}\begin{aligned}
	H_D^{(j)} = \varepsilon_j\sum_{q,\sigma}\tau_{q,\sigma} + \frac{1}{2}\sum_{{\bf q}}J_{{\bf q}, {\bf q}}S_d^z\left(\hat n_{{\bf q}, \uparrow} - \hat n_{{\bf q}, \downarrow}\right) - \frac{1}{2}\sum_{{\bf q}}W_{\bf q}\left(\hat n_{{\bf q}, \uparrow} - \hat n_{{\bf q}, \downarrow}\right)^2~,
\end{aligned}\end{equation}
where \(\tau = \hat n - 1/2\) and \(W_{{\bf q}}\) is a shorthand for \(W_{{\bf q},{\bf q},{\bf q},{\bf q}}\). The three terms, respectively, are the kinetic energy of the momentum states on the isoenergetic shell that we are decoupling, the spin-correlation energy between the impurity spin and the spins formed by these momentum states and, finally, the local correlation energy associated with these states arising from the \(W\) term. The off-diagonal part of the Hamiltonian on the other hand leads to scattering in the states \(\left\{ {\bf q} \right\} \). We now list these terms, classified by the coupling they originate from.

@@

\section{Tiling formalism: Mapping from the auxiliary model to the tiled lattice model}
The eigenstates \(\ket{\Psi_{s}}\) (\(s = \left({\bf k}, n\right)\)) of the lattice Hamiltonians obtained using  the tiling procedure outlined in the main text enjoy a {\it many-body} Bloch's theorem~[55],
%\cite{stoyanova}, 
because the tiling procedure restores the translation symmetry of the Hamiltonian. This means that the {\it local} eigenstates \(\ket{\psi_n\left({\bf r}_d\right)}\) (with the impurity located at an arbitrary position \({\bf r}_d\)) of the unit cell auxiliary model Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) can be used to construct eigenstates of the lattice Hamiltonian. The index \(n(=0,1,\ldots)\) in the subscript indicates that it is the \(n^\text{th}\) eigenstate of the auxiliary model.

The state \(\ket{\psi_n\left({\bf r}_d\right)}\) does not specify the position of the zeroth site, because the unit cell Hamiltonian \(\mathcal{H}_\text{aux}({\bf r}_d)\) itself has been averaged over \(\mathcal{Z}\) zeroth sites. Accordingly, we can express the averaged eigenstate \(\ket{\psi_n\left({\bf r}_d\right)}\) as
\begin{equation}\begin{aligned}
	\ket{\psi_n\left({\bf r}_d\right)} = \frac{1}{\sqrt\mathcal{Z}}\sum_{{\bf z} \in \text{NN}({\bf r}_d)}\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(\ket{\psi_n\left({\bf r}_d, {\bf z}\right)}\) is an auxiliary model eigenstate with the impurity and zeroth sites placed at \({\bf r}_d\) and \({\bf z}\). With this in mind, the following unnormalised combination of the auxiliary model eigenstates satisfies a many-particle equivalent of Bloch's theorem~[55]:
%\cite{stoyanova}:
\begin{equation}\begin{aligned}\label{eigenstateProposal1}
	\ket{\Psi_{s}} \equiv \ket{\Psi_{{\bf k}, n}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_{n}\left({\bf r}_d, {\bf z}\right)}~,
\end{aligned}\end{equation}
where \(N\) is the total number of lattice sites and \({\bf r}_d\) is summed over all lattice spacings. The set of \(n=0\) states form the lowest band in the spectrum of the lattice, while higher values of \(n\) produce the more energetic bands. The ground state \(s = s_0\) is obtained by setting \({\bf k}\) and \(n\) to 0:
\begin{equation}\begin{aligned}\label{groundstateProposal1}
	\ket{\Psi_\text{gs}} \equiv \ket{\Psi_{s_0}} &= \frac{1}{\sqrt N}\sum_{{\bf r}_d} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d\right)} = \frac{1}{\sqrt{\mathcal{Z} N}}\sum_{{\bf r}_d}\sum_{{\bf z} \in \text{NN}({\bf r}_d)} e^{i {\bf k}\cdot{\bf r}_d} \ket{\psi_\text{gs}\left({\bf r}_d, {\bf z}\right)}
\end{aligned}\end{equation}

The retarded time-domain lattice \(k-\)space Greens function is defined as
	\tilde G({\bf K}\sigma; t) = -i\theta(t) \braket{\Psi_\text{gs} | \left\{ c_{{\bf K}\sigma}(t), c^\dagger_{{\bf K}\sigma} \right\} | \Psi_\text{gs}}~.
where the bulk Hamiltonian \(H_\text{tiled}\) leads to the dynamics of the annihilation operators at time \(t\): 
\begin{equation}\begin{aligned}\label{heisenberg}
	c_{{\bf K}\sigma}(t) = e^{it H_\text{tiled} }c_{{\bf K}\sigma}e^{-i t H_\text{tiled}}~.
We now proceed to simplify one of the terms of the anticommutator (for simplicity of notation):
\begin{equation}\begin{aligned}\label{greensFunction1}
	&\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{N^2}\sum_{\vec r,\vec \Delta}e^{-i{\bf K}_0\cdot\vec\Delta}\braket{\psi_{0}(\vec r+\vec \Delta) | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~.
Using eq.~\ref{eigenstateProposal1} and the identity resolution \(1 = \sum_s\ket{\Psi_s}\bra{\Psi_s}\), eq.~\ref{greensFunction1} becomes
	\begin{equation}\begin{aligned}\label{greensfunction2}
		\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{N^2}\sum_s\sum_{\vec r,\vec \Delta}\sum_{\vec r^\prime,\vec \Delta^\prime}e^{-i\vec k_0\cdot\vec\Delta} e^{i\vec k\cdot\vec\Delta^\prime}\braket{\psi_{0}(\vec r+\vec \Delta) | c_{{\bf K}\sigma}(t) \ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~.
In order to bring this expression closer to the form of an auxiliary model Greens function, we
\begin{itemize}
    \item use the relation $\ket{\psi(\vec r + \Delta)} = T^\dagger(\vec \Delta)\ket{\psi(\vec r)}$, where \(T^\dagger(\vec \Delta)\) translates all lattice sites by the vector \(\vec\Delta\),
    \item use the property $T\left(\vec a\right) c({\bf K}) T^\dagger\left(\vec a\right) = e^{-i{\bf K}\cdot\vec a}c({\bf K})$,
    \item make the substitution \(\vec \Delta^\prime \to \vec \Delta^\prime + \vec \Delta\).
\end{itemize}
This leads to the expression
\begin{equation}\begin{aligned}\label{greensfunction3}
		\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{N}\sum_n\sum_{\vec r,\vec r^\prime,\vec \Delta^\prime} e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\braket{\psi_{0}(\vec r) | c_{{\bf K}\sigma}(t) \ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~,
where the sum over \(s=(\vec k, n)\) has been reduced to a sum over the auxiliary model eigenstate index \(n\) because of the Kronecker delta \(\delta\left(\vec k_0 + {\bf K} - \vec k\right)\). This can be further simplified by splitting the sum over \(\vec\Delta^\prime\) into positive and negative parts and then making the transformation \(\vec r^\prime \to \vec r^\prime + \vec \Delta^\prime\):
	\sum_{\vec r^\prime,\vec \Delta^\prime} e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} &= \frac{1}{2}\sum_{\vec r^\prime,\vec \Delta^\prime}\left[ e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} + e^{-i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime-\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)}\right] \\
																																												   &= \frac{1}{2}\sum_{\vec r^\prime,\vec \Delta^\prime}\left[ e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} + e^{-i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime)}\bra{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\right] \\
For each pair of \(\vec r^\prime\) and \(\vec\Delta^\prime\), the term within the box brackets has the form of a two-level Hamiltonian between the states \(\ket{\psi_n(\vec r^\prime)}\) and \(\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\), with a tunnelling amplitude \(e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\). The term can therefore be written in the eigenbasis of this Hamiltonian:
	\sum_{\vec r^\prime,\vec \Delta^\prime} e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime+\vec \Delta^\prime)}\bra{\psi_n(\vec r^\prime)} &=\frac{1}{2}\sum_{\vec r^\prime,\vec \Delta^\prime}\left[ \ket{\chi_n^+(\vec r^\prime,\vec\Delta^\prime)}\bra{\chi_n^+(\vec r^\prime,\vec\Delta^\prime)} - \ket{\chi_n^-(\vec r^\prime,\vec\Delta^\prime)}\bra{\chi_n^-(\vec r^\prime,\vec\Delta^\prime)}\right],
where \(\ket{\chi_n^\pm(\vec r^\prime,\vec\Delta^\prime)} = \frac{1}{\sqrt 2}\left[\ket{\psi_n(\vec r^\prime)} \pm e^{i\left( \vec k_0 + {\bf K} \right) \cdot\vec\Delta^\prime}\ket{\psi_n(\vec r^\prime + \vec\Delta^\prime)}\right] \) are the eigenvectors of the tunnelling Hamiltonian with eigenvalues \(\pm 1\) respectively. With this basis transformation, we can rewrite eq.~\ref{greensfunction3} as
		\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{2N}\sum_n\sum_{\vec r,\vec r^\prime,\vec \Delta^\prime} \braket{\psi_{0}(\vec r) | c_{{\bf K}\sigma}(t) \left[ \ket{\chi_n^+(\vec r^\prime,\vec\Delta^\prime)}\bra{\chi_n^+(\vec r^\prime,\vec\Delta^\prime)} - \ket{\chi_n^-(\vec r^\prime,\vec\Delta^\prime)}\bra{\chi_n^-(\vec r^\prime,\vec\Delta^\prime)}\right] c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)}~.

In the present work, we consider only the $\vec r^\prime = \vec r,~\vec\Delta^\prime=0$ component. These terms represent those contributions to the total Greens function that arise from excitations that start and end at a specific auxiliary model (at \(\vec r\)), and also evolve dynamically within the same auxiliary model. These terms are therefore exactly equal to the auxiliary model Greens function at position \(\vec r\), and are the most dominant contribution due to the localised nature of the impurity model.

Restricting ourselves to just the single auxiliary model contributions gives
\begin{equation}\begin{aligned}
		\braket{\Psi_\text{gs} | c_{{\bf K}\sigma}(t) c^\dagger_{{\bf K}\sigma} | \Psi_\text{gs}} = \frac{1}{N}\sum_n\sum_{\vec r} \braket{\psi_{0}(\vec r) | c_{{\bf K}\sigma}(t) \ket{\psi_n(\vec r)}\bra{\psi_n(\vec r)} c^\dagger_{{\bf K}\sigma} | \psi_{0}(\vec r)} ~.
\end{aligned}\end{equation}
We first consider more carefully the transition operator \(\mathcal{T}_{{\bf K}\sigma} = c_{{\bf K}\sigma}\) for the 1-particle excitation giving rise to the above Greens function. Within our auxiliary model approach, gapless excitations within the lattice model are represented by gapless excitations of the impurity site, specifically those that screen the impurity site and form the local Fermi liquid. As a result, the uncoordinated \(\mathcal{T}-\)matrix for the lattice model must be replaced by a combined \(\mathcal{T}-\)matrix within the impurity model that captures those gapless excitations that occur in connection with the impurity, and projects out the uncorrelated excitations that take place even when the impurity site is decoupled from the bath.

In order to construct this auxiliary model \(\mathcal{T}-\)matrix, we note that the impurity site can have both spin and charge excitations. Considering both excitations, the modified \(\mathcal{T}-\)matrix that constructs \(k-\)space excitations in correlation with the impurity site are
\begin{equation}\begin{aligned}\label{tmatrix}
	\mathcal{T}_{{\bf K}\sigma} = c_{{\bf K}\sigma}\left(\sum_{\sigma^\prime}c^\dagger_{d\sigma} + \text{h.c.}\right) + c_{{\bf K}\sigma}\left(S_d^+ + \text{h.c.}\right)~,
\end{aligned}\end{equation}
leading to the updated expression for the complete Greens function:
\begin{equation}\begin{aligned}
	\tilde G({\bf K}\sigma; t) = -i\theta(t)\frac{1}{N}\sum_n\sum_{\vec r} \braket{\psi_{0}(\vec r) | \left[\mathcal{T}_{{\bf K}\sigma}(t) \ket{\psi_n(\vec r)}\bra{\psi_n(\vec r)} \mathcal{T}_{{\bf K}\sigma}^\dagger + \mathcal{T}^\dagger_{{\bf K}\sigma} \ket{\psi_n(\vec r)}\bra{\psi_n(\vec r)} \mathcal{T}_{{\bf K}\sigma}(t)\right] | \psi_{0}(\vec r)} ~.
\end{aligned}\end{equation}

@@

\section{Finite-Frequency Optical Response Through the Pseudogap}
\begin{figure}[!ht]
	\includegraphics[width=0.5\textwidth]{sigma_in.pdf}
	\caption{Optical conductivity $\sigma(\omega)$ for excitations of the impurity model, approximated through the product of the carrier density (taken from the density of states $A(\omega)$) and the lifetime $\tau(\omega)$ (obtained from the inverse scattering rate $-1/\Sigma^{\prime\prime}(\omega)$, where $\Sigma^{\prime\prime}$ is the imaginary part of the self-energy). The Fermi liquid (blue) shows a sharp Drude peak at $\omega=0$ due to the presence of long-lived quasiparticles. The PG (green, red, orange) shows a shifted Drude peak at a non-zero frequency which is reminiscent of the mid-infrared (MIR) peak seen experimentally in the cuprates. Inset shows the robustness of the position of the MIR peak through the PG, indicating that it's a universal feature within the PG of our model.}
	\label{optical}
\end{figure}
To further probe the breakdown of coherent transport in the pseudogap (PG) regime, we examine the product of the impurity spectral function $A(\omega)$ and the scattering lifetime $\tau(\omega)$, which approximates the optical conductivity $\sigma(\omega)$ via a Drude-type relation. As shown in the main panel of Fig.~\ref{optical}, this quantity exhibits a sharp Drude peak at $\omega = 0$ in the Fermi liquid (FL) regime (blue), consistent with long-lived quasiparticles. In stark contrast, the PG regime (green, red, orange) shows a pronounced suppression of low-frequency conductivity and the emergence of a finite-frequency peak centered near $\omega/D \sim 0.3{-}0.4$ (where $D$ is the kinetic energy bandwidth).

Strikingly, this peak shifts very little across the PG regime (see inset), despite a deepening spectral gap and diverging self-energy. Its position lies well below the Mott gap ($\sim 2.5D$), suggesting that it represents an emergent mid-infrared (MIR) excitation intrinsic to the Mott metal characterising the PG. This feature closely mirrors the experimentally observed MIR peak in high-T$_\text{c}$ cuprates~\cite{Herr1987,Orenstein1987,Uchida1991}, which has been variously attributed to excitations of electrons bound to vacancies~\cite{Thomas1992Optical}, or the dressing of holes by spin excitations~\cite{Dagotto1994RMP}. Its robust energy scale and insensitivity to increasing incoherence imply a universality persisting through the PG, likely rooted in short-range entanglement between deconfined doublons and holons excitations of the Mott metal. This is further evidence that the PG metal is not simply a degraded Fermi liquid, but a distinct phase whose optical response is driven by emergent degrees of freedom.

% To further probe the breakdown of coherent transport in the pseudogap (PG) regime, we examine the product of the impurity spectral function $A(\omega)$$ and the scattering lifetime $\tau(\omega)$$, which approximates the optical conductivity $\sigma(\omega)$ via a Drude-type relation. As shown in Fig.(REF)[X], this quantity exhibits a sharp Drude peak at $\omega = 0$ in the Fermi liquid (FL) regime, consistent with long-lived quasiparticles. In stark contrast, the PG regime (dashed curves) shows a pronounced suppression of low-frequency conductivity and the emergence of a finite-frequency peak centered near $\omega/D \sim 0.3{-}0.4$ (where $D$ is the kinetic energy bandwidth). Strikingly, this peak shifts very little across the PG regime, despite a deepening spectral gap and diverging self-energy. Its position lies well below the Mott gap ($\sim 2.5D$), suggesting that it represents an emergent mid-infrared excitation intrinsic to the PG–Mott crossover. This feature closely mirrors the experimentally observed mid-IR peak in high-T$_{\textrm{c}}$ cuprates (CITE), which has been variously attributed to preformed pairs (CITE), bound holon–doublon states (CITE), or emergent collective modes (CITE). Its robust energy scale and insensitivity to increasing incoherence imply a hidden order or resonance persisting across the PG regime, likely rooted in short-range entanglement between deconfined doublons and holons. Thus, the PG metal is not simply a degraded Fermi liquid, but a distinct dynamical phase whose optical response encodes the fingerprints of emergent quantum degrees of freedom.

@@

\section{Additional Results on the Fermi Liquid \& Pseudogap-Mott metal Phases}
We present additional figures of the spin (Fig.\ref{spinCorr}) and charge (Fig.\ref{cfnode}) correlations in the impurity model, the impurity spectral function (Fig.\ref{specfunc}), the spin correlations (Fig.\ref{tiledSF}), entanglement entropy (Fig.\ref{tiledEntanglement} (upper panel)) and mutual function (Fig.\ref{tiledEntanglement} (lower panel)) of the tiled model in the pseudogap phase. All figures presented here are for a $77\times 77$ {\bf k}-space Brillouin zone grid.
\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.32\textwidth]{SF-500-1.pdf}
	\includegraphics[width=0.32\textwidth]{SF-500-3.pdf}
	\includegraphics[width=0.32\textwidth]{SF-500-4.pdf}
	\caption{\(k-\)space distribution of spin-spin correlation \(\chi_s(d,\vec k)\) between the impurity spin and momentum states in the conduction bath. The first shows a mostly uniform distribution of correlations on the Fermi surface for $W=0$. The second plot shows the same inside the PG: the antinodal region no longer participates in Kondo screening. The last plot represents the nodal metal emerging at the critical point, where the entirety of Kondo screening is carried out by a small neighbourhood about the nodes.}
	\label{spinCorr}
\end{figure}

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.4\linewidth]{Ad_zoomout_77-1500.pdf}
    \caption{Impurity Spectral function in the FL and PG phases. The evolution of the central peak from Kondo resonance to pseudogap is accompanied by dynamical spectral weight transfer to the Hubbard side bands at finite frequencies $\omega\simeq \pm 3$ (in units of the bandwidth).}
    \label{specfunc}
\end{figure}

@@

%\bibliography{tilingProject}
%\bibliographystyle{unsrt}

\begin{thebibliography}{99}
\bibliographystyle{unsrt}

\bibitem{santanukagome}
Santanu Pal, Anirban Mukherjee, and Siddhartha Lal.
\newblock Correlated spin liquids in the quantum kagome antiferromagnet at
  finite field: a renormalization group analysis.
\newblock {\em New Journal of Physics}, 21(2):023019, feb 2019.

\bibitem{1dhubjhep1}
Anirban Mukherjee, Siddhartha Patra, and Siddhartha Lal.
\newblock Fermionic criticality is shaped by fermi surface topology: a case
  study of the tomonaga-luttinger liquid.
\newblock {\em Journal of High Energy Physics}, 2021(4):148, Apr 2021.

\bibitem{anirbanmott1}
Anirban Mukherjee and Siddhartha Lal.
\newblock Scaling theory for mott{\textendash}hubbard transitions: I. t = 0
  phase diagram of the 1/2-filled hubbard model.
\newblock {\em New Journal of Physics}, 22(6):063007, jun 2020.

\bibitem{siddharthacpi}
Siddhartha Patra and Siddhartha Lal.
\newblock Origin of topological order in a cooper-pair insulator.
\newblock {\em Phys. Rev. B}, 104:144514, Oct 2021.

\bibitem{anirban_kondo}
Anirban Mukherjee, Abhirup Mukherjee, N.~S. Vidhyadhiraja, A.~Taraphder, and
  Siddhartha Lal.
\newblock Unveiling the kondo cloud: Unitary renormalization-group study of the
  kondo model.
\newblock {\em Phys. Rev. B}, 105:085119, Feb 2022.

>>>>

%\documentclass{article}
\documentclass[
preprint,
superscriptaddress,
groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
%aps,superscriptaddress,
aps,pra,
%rmp,
%prstab,
%prstper,
%floatfix,
10pt
]{revtex4-2}
%\usepackage{amsmath,amssymb,bm,graphicx,braket}
\usepackage[margin=1.5cm]{geometry}
%\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{graphicx,braket,color,bm,amsmath,amssymb,braket,placeins}
% \graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}
\usepackage{natbib}

@@

\makeatletter
\apptocmd{\thebibliography}{\global\c@NAT@ctr 80\relax}{}{}
\makeatother
\begin{document}
\title{Supplementary Materials for ``Mott Criticality as the Confinement Transition of a Pseudogap-Mott Metal"}

\author{Abhirup Mukherjee}
\affiliation{
 Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}
\author{S. R. Hassan}
\affiliation{The Institute of Mathematical Sciences, C.I.T. Campus, Chennai 600 113, India}
\author{Anamitra Mukherjee}
\affiliation{School of Physical Sciences, National Institute of Science, Education and Research, HBNI, Jatni 752050, India}
\author{N. S. Vidhyadhiraja}
\affiliation{Theoretical Sciences Unit, Jawaharlal Nehru Center for Advanced Scientific Research, Jakkur, Bengaluru 560064, India}
\author{A. Taraphder}
\affiliation{Department of Physics, Indian Institute of Technology Kharagpur, Kharagpur 721302, India}
\author{Siddhartha Lal}
\affiliation{
 Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}

@@

\section{Derivation of unitary RG equations for the lattice-embedded impurity model}
At any given step \(j\) of the RG procedure, we decouple the states \(\left\{ {\bf q} \right\} \) on the isoenergetic surface of energy \(\varepsilon_j\). The diagonal Hamiltonian \(H_D\) for this step consists of all terms that do not change the occupancy of the states \(\left\{{\bf q}\right\}\):
\begin{equation}\begin{aligned}
	H_D^{(j)} = \varepsilon_j\sum_{q,\sigma}\tau_{q,\sigma} + \frac{1}{2}\sum_{{\bf q}}J_{{\bf q}, {\bf q}}S_d^z\left(\hat n_{{\bf q}, \uparrow} - \hat n_{{\bf q}, \downarrow}\right) - \frac{1}{2}\sum_{{\bf q}}W_{\bf q}\left(\hat n_{{\bf q}, \uparrow} - \hat n_{{\bf q}, \downarrow}\right)^2~,
\end{aligned}\end{equation}
where \(\tau = \hat n - 1/2\) and \(W_{{\bf q}}\) is a shorthand for \(W_{{\bf q},{\bf q},{\bf q},{\bf q}}\). The three terms, respectively, are the kinetic energy of the momentum states on the isoenergetic shell that we are decoupling, the spin-correlation energy between the impurity spin and the spins formed by these momentum states and, finally, the local correlation energy associated with these states arising from the \(W\) term. The off-diagonal part of the Hamiltonian on the other hand leads to scattering in the states \(\left\{ {\bf q} \right\} \). We now list these terms, classified by the coupling they originate from.

@@

@@

@@

\section{Additional Results on the Fermi Liquid \& Pseudogap-Mott metal Phases}
We present additional figures of the spin (Fig.\ref{spinCorr}) and charge (Fig.\ref{cfnode}) correlations in the impurity model, the impurity spectral function (Fig.\ref{specfunc}), the spin correlations (Fig.\ref{tiledSF}), entanglement entropy (Fig.\ref{tiledEntanglement} (upper panel)) and mutual function (Fig.\ref{tiledEntanglement} (lower panel)) of the tiled model in the pseudogap phase.
\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.32\textwidth]{SF-500-1.pdf}
	\includegraphics[width=0.32\textwidth]{SF-500-3.pdf}
	\includegraphics[width=0.32\textwidth]{SF-500-4.pdf}
	\caption{\(k-\)space distribution of spin-spin correlation \(\chi_s(d,\vec k)\) between the impurity spin and momentum states in the conduction bath. The first shows a mostly uniform distribution of correlations on the Fermi surface for $W=0$. The second plot shows the same inside the PG: the antinodal region no longer participates in Kondo screening. The last plot represents the nodal metal emerging at the critical point, where the entirety of Kondo screening is carried out by a small neighbourhood about the nodes.}
	\label{spinCorr}
\end{figure}

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.4\linewidth]{figures/Ad_zoomout_77-1500.pdf}
    \caption{Impurity Spectral function in the FL and PG phases. The evolution of the central peak from Kondo resonance to pseudogap is accompanied by dynamical spectral weight transfer to the Hubbard side bands at finite frequencies $\omega\simeq \pm 3$ (in units of the bandwidth).}
    \label{specfunc}
\end{figure}

@@

\bibliography{tilingProject}
\bibliographystyle{unsrt}
